Scipion Tailored for TEM Facilities: Image Processing in Streaming

### 2018-02-19 2.overall.tex
going home

The workflow starts two independent protocols in parallel. The first protocol estimates the camera gain at each pixel for a few  movies \citep{sorzano2018}. % In the CNB setup the microscope camera is calibrated and internally performs the gain correction before writing the movies so the estimated gain image should be almost constant and equal to 1. 
The output of this protocol serves as verification of the validity of the experimental gain image for the whole data set. The second protocol performs Scipion direct detector movie alignment. Currently \scipion offers  four  methods  for  DDD  movie  alignment: motioncorr2 (Li et al., 2013), correlation (an unpublished CPU version of the motioncorr2 method developed in Xmipp), unblur (Grant and Grigorieff, 2015) and optical flow (Abrishami et al., 2015). Within \scipion, movie frames can be aligned using any one of these approaches or any combination. Following the standard processing workflow, the resulting micrographs are screened to eliminate those with excessive astigmatism or drift. Scipion provides three CTF estimation methods:
CTFFIND (Rohou and Grigorieff, 2015), gCTF (Zhang, 2016) and Xmipp CTF estimation (Sorzano et al., 2007; Vargas et al., 2013). At this point the user may compute the CTFs using several algorithms and compare them with the protocol \protocol{CTF consensus} that report up to which frequency two estimation of the same micrograph are equivalent. Further automatic filtering based on CTF is available through the protocol \protocol{CTF filtering} that selects micrographs based on the astigmatism, resolution and defocus. Another criteria for CTF selection is iciness, that is,  the ratio of the intensity in the resolution band between 3.5 and 3.9 \AA, and between 6 and 30 \AA, which gives a good estimate of the ice crystal content of the vitreous specimen. [**ROB: do we have iceness implemented?]

Up to this point the image processing pipeline is quite solid and may be executed in streaming without user intervention. The next step is to select (pick) particles. We have recently ported all the  different options available at \scipion  to streaming. In this way,  particles may be selected as soon as any new micrograph is available. We note here that some of the ported algorithms require a training step in which either user intervention or information from a previous selection is need. For those users that execute several picking algorithms, the protocol \protocol{picking consensus} may be useful. \protocol{picking consensus}  calculates the agreement between different particle-picking algorithms and produces a list with the particle selected by most of them.

Two new protocols are under development and will join the described  pipeline in a near future. The first one will perform 2D classification and the second one will create a initial 3D map based on this classes. A first version of these protocols is available in the \scipion developement version.

As commented in section \ref{overall}, to complement the streaming process we have developed a set of monitors and report possibilities. By default \protocol{system}, \protocol{CTF} and \protocol{report} monitors are executed. \protocol{System monitor} reports on  CPU and GPU utilization, memory consumption and disk I/O. \protocol{CTF monitor} plots the CTF defocus and astigmatism  and it will raise an alarm if these parameters are not within a given range. Finally the \protocol{report} protocol pushes all the created plots and summaries to a web server 
so this information can be accessed from outside.

>>>>

The workflow starts two independent protocols in parallel. The first protocol estimates the camera gain at each pixel for a few  movies \citep{sorzano2018}. % In the CNB setup the microscope camera is calibrated and internally performs the gain correction before writing the movies so the estimated gain image should be almost constant and equal to 1. 
The output of this protocol serves as verification of the validity of the experimental gain image for the whole data set. The second protocol performs Scipion direct detector movie alignment. Currently \scipion offers  four  methods  for  DDD  movie  alignment: motioncorr2 (Li et al., 2013), correlation (an unpublished CPU version of the motioncorr2 method developed in Xmipp), unblur (Grant and Grigorieff, 2015) and optical flow (Abrishami et al., 2015). Within \scipion, movie frames can be aligned using any one of these approaches or any combination. Following the standard processing workflow, the resulting micrographs are screened to eliminate those with excessive astigmatism or drift. Scipion provides three CTF estimation methods:
CTFFIND (Rohou and Grigorieff, 2015), gCTF (Zhang, 2016) and Xmipp CTF estimation (Sorzano et al., 2007; Vargas et al., 2013). At this point the user may compute the CTFs using several algorithms and compare them with the protocol \protocol{CTF consensus} that report up to which frequency two estimation of the same micrograph are equivalent. Further automatic filtering based on CTF is available through the protocol \protocol{CTF filtering} that selects micrographs based on the astigmatism, resolution and defocus. Another criteria for CTF selection is iciness, that is,  the ratio of the intensity in the resolution band between 3.5 and 3.9 \AA, and between 6 and 30 \AA, which gives a good estimate of the ice crystal content of the vitreous specimen.

Up to this point the image processing pipeline is quite solid and may be executed in streaming without user intervention. The next step is to select (pick) particles. We have recently ported the eight different options available at \scipion  to streaming. In this way,  particles may be selected as soon as any new micrograph is available. We note here that some of the ported algorithms require a training step in which either user intervention or information from a previous selection is need. For those users that execute several picking algorithms, the protocol \protocol{picking consensus} may be useful. \protocol{picking consensus}  calculates the agreement between different particle-picking algorithms and produces a list with the particle selected by most of them.

Two new streaming protocols are under development and will join the described  pipeline in a near future. The first one will perform 2D classification and the second one will create a initial 3D map based on this classes. A first version of these protocols is available in the \scipion developers version.

As commented in section \ref{overall}, to monitor, chart and create alerts  we have developed a set of metaprotocols. By default \protocol{system}, \protocol{CTF} and \protocol{report monitors} are executed. \protocol{System monitor} reports on  CPU and GPU utilization, memory consumption and, disk I/O. \protocol{CTF monitor} plots the CTF defocus and astigmatism  and it will raise an alarm if these parameters are not within a given range. Finally the \protocol{report} protocol pushes all the created plots and summaries to a web server 
so this information can be accessed by users who are not physically in the microscope room.
